---
title: "Analysis of Cumulative Emissions and Odor Activity Values from pig slurry"
author: "Flavia"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE}
# Load required libraries
library(emmeans)
library(multcompView)
library(multcomp)
library(readr)
library(dplyr)
library(ggplot2)
library(olsrr)
library(knitr)

# Set chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

# Introduction

This analysis investigates the effect of acid dose added to pig slurry (0, 2.9, 5.3, and 10.5 kg per ton of slurry) on cumulative emissions of non-methane volatile organic compounds (VOCs), ammonia (NH3), and odor activity values (OAV), measured using dynamic flux chambers (DFC). Additionally, emissions (only from untreated slurry, i.e. dose 0) were measured from a big plot where slurry was machine-applied (Machine plot), and from small plots where it was manually applied. We want to compare how application method influences our results.

# Cumulative NMVOC Emissions

## Data Loading and Preparation

```{r load-voc-data}
# Load cumulative VOC data
cum.voc <- read_csv("../Flavia_VOC_DFC_data/cum.voc.emis.csv")

# Show data structure
kable(head(cum.voc), caption = "First 6 rows of cumulative VOC data")
```

```{r prepare-voc-data}
# Filter out machine plot treatment
filtered_data_voc <- cum.voc[!cum.voc$group == 'Machine plot', ]

# Assign dose values based on group names
doses <- c(`No acid` = 0, `Low acid` = 2.9, `Medium acid` = 5.3, `High acid` = 10.5)
filtered_data_voc$dose <- doses[filtered_data_voc$group]

# Ensure dose is numeric
filtered_data_voc$dose <- as.numeric(filtered_data_voc$dose)

# Summary of prepared data
cat("Sample sizes by dose:\n")
table(filtered_data_voc$dose)
```

## Dose-Response Analysis

```{r voc-dose-model}
# Fit linear model
model <- lm(total_cum ~ dose, data = filtered_data_voc)

# Model summary
summary(model)

# Detailed regression output
ols_regress(model)
```

### Model Diagnostics

```{r voc-diagnostics}
# Levene's test for homoscedasticity
res_abs <- abs(residuals(model))
levene <- lm(res_abs ~ dose, data = filtered_data_voc)
levene_result <- anova(levene)

# Shapiro-Wilk test for normality
shapiro_result <- shapiro.test(residuals(model))

# Display results
cat("Levene's Test for Homoscedasticity:\n")
print(levene_result)
cat("\nShapiro-Wilk Test for Normality:\n")
print(shapiro_result)
```

### Visualization

```{r voc-dose-plot, fig.cap="Cumulative VOC emissions by acid dose"}
p1 <- ggplot(filtered_data_voc, aes(x = dose, y = total_cum, color = as.factor(dose))) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, color = "black", linetype = "dashed") +
  labs(title = "Cumulative VOC Emissions by Acid Dose",
       x = "Acid Dose",
       y = expression("Cumulative VOC emissions"~(mg~m^{-2})),
       color = "Dose") +
  theme_minimal() +
  theme(legend.position = "right")

print(p1)
```

## Machine vs Manual Application

```{r voc-machine-manual}
# Filter for comparison
filtered_data_2 <- cum.voc[cum.voc$group %in% c('Machine plot', 'No acid'), ]

# Fit model
model_2 <- lm(total_cum ~ group, data = filtered_data_2)

# ANOVA
anova_result_2 <- anova(model_2)
print(anova_result_2)

# Diagnostics
shapiro_2 <- shapiro.test(model_2$residuals)
filtered_data_2$res_abs <- abs(model_2$residuals)
levene_2 <- anova(lm(res_abs ~ group, data = filtered_data_2))

cat("Normality test p-value:", shapiro_2$p.value, "\n")
cat("Homoscedasticity test p-value:", levene_2$`Pr(>F)`[1], "\n")

# Extract means
means_2 <- emmeans(model_2, ~ group)
print(means_2)
```

```{r voc-application-plot, fig.cap="VOC emissions: Machine vs Manual application"}
p2 <- ggplot(filtered_data_2, aes(x = group, y = total_cum, color = group)) +
  geom_point(size = 3, alpha = 0.7) +
  stat_summary(fun = mean, geom = "crossbar", width = 0.5, color = "black") +
  labs(title = "VOC Emissions: Application Method Comparison",
       x = "Application Method",
       y = expression("Cumulative VOC emissions"~(mg~m^{-2})),
       color = "Group") +
  theme_minimal() +
  theme(legend.position = "none")

print(p2)
```

# Cumulative NH3 Emissions

## Data Loading and Analysis

```{r load-nh3-data}
# Load NH3 data
cum_NH3 <- read_delim("cum.NH3.csv", delim = ";", 
                      escape_double = FALSE, trim_ws = TRUE)

# Show data structure
kable(head(cum_NH3), caption = "First 6 rows of cumulative NH₃ data")
```

```{r prepare-nh3-data}
# Filter and prepare data
filtered_data_NH3 <- cum_NH3[!cum_NH3$treatment == 'Machine plot DFC', ]

# Assign dose values based on treatment names
doses_nh3 <- c(`No acid DFC` = 0, `Low acid DFC` = 2.9, 
               `Medium acid DFC` = 5.3, `High acid DFC` = 10.5)
filtered_data_NH3$dose <- doses_nh3[filtered_data_NH3$treatment]

# Ensure dose is numeric
filtered_data_NH3$dose <- as.numeric(filtered_data_NH3$dose)

# Sample sizes
cat("Sample sizes by dose:\n")
table(filtered_data_NH3$dose)
```

## Dose-Response Analysis

```{r nh3-dose-model}
# Fit model
model_nh3 <- lm(cum.emis.TAN ~ dose, data = filtered_data_NH3)

# Results
summary(model_nh3)
ols_regress(model_nh3)
```

### Model Diagnostics

```{r nh3-diagnostics}
# Diagnostics
res_abs_nh3 <- abs(residuals(model_nh3))
levene_nh3 <- anova(lm(res_abs_nh3 ~ dose, data = filtered_data_NH3))
shapiro_nh3 <- shapiro.test(residuals(model_nh3))

cat("Levene's Test p-value:", levene_nh3$`Pr(>F)`[1], "\n")
cat("Shapiro-Wilk Test p-value:", shapiro_nh3$p.value, "\n")
```

```{r nh3-dose-plot, fig.cap="Cumulative NH₃ emissions by acid dose"}
p3 <- ggplot(filtered_data_NH3, aes(x = dose, y = cum.emis.TAN, color = as.factor(dose))) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, color = "black", linetype = "dashed") +
  labs(title = "Cumulative NH₃ Emissions by Acid Dose",
       x = "Acid Dose",
       y = "Cumulative NH₃ emissions (% TAN)",
       color = "Dose") +
  theme_minimal() +
  theme(legend.position = "right")

print(p3)
```

## Machine vs Manual Application (NH₃)

```{r nh3-machine-manual}
# Filter and analyze
filtered_data_3 <- cum_NH3[cum_NH3$treatment %in% c('Machine plot DFC', 'No acid DFC'), ]

model_3 <- lm(cum.emis.TAN ~ treatment, data = filtered_data_3)
anova_3 <- anova(model_3)
print(anova_3)

# Diagnostics
shapiro_3 <- shapiro.test(model_3$residuals)
filtered_data_3$res_abs <- abs(model_3$residuals)
levene_3 <- anova(lm(res_abs ~ treatment, data = filtered_data_3))

cat("Normality test p-value:", shapiro_3$p.value, "\n")
cat("Homoscedasticity test p-value:", levene_3$`Pr(>F)`[1], "\n")

# Means
means_3 <- emmeans(model_3, ~ treatment)
print(means_3)
```

```{r nh3-application-plot, fig.cap="NH₃ emissions: Machine vs Manual application"}
p4 <- ggplot(filtered_data_3, aes(x = treatment, y = cum.emis.TAN, color = treatment)) +
  geom_point(size = 3, alpha = 0.7) +
  stat_summary(fun = mean, geom = "crossbar", width = 0.5, color = "black") +
  labs(title = "NH₃ Emissions: Application Method Comparison",
       x = "Treatment",
       y = "Cumulative NH₃ emissions (% TAN)",
       color = "Treatment") +
  theme_minimal() +
  theme(legend.position = "none")

print(p4)
```

# Odor Activity Values (OAV)

## Data Loading and Temporal Analysis

```{r load-oav-data}
# Load OAV data
OAV <- read_csv("../Flavia_VOC_DFC_data/OAV.txt")

# Filter data
filtered_data_OAV <- OAV[!OAV$group == 'Machine plot', ]

# Assign dose values based on group names
doses_oav <- c(`No acid` = 0, `Low acid` = 2.9, `Medium acid` = 5.3, `High acid` = 10.5)
filtered_data_OAV$dose <- doses_oav[filtered_data_OAV$group]

# Show data structure
kable(head(filtered_data_OAV), caption = "First 6 rows of OAV data")
```

## Slope and Intercept Analysis

```{r oav-slope-intercept}
# Calculate intercept and slope for each valve
fit <- filtered_data_OAV %>% 
  group_by(valve) %>% 
  summarise(
    intercept = coef(lm(total_OAV ~ poly(elapsed_time, 1, raw = TRUE)))[1],
    slope = coef(lm(total_OAV ~ poly(elapsed_time, 1, raw = TRUE)))[2],
    .groups = 'drop'
  )

# Join with dose information
fit.dose <- left_join(fit, 
                      filtered_data_OAV %>% 
                        select(valve, dose) %>% 
                        distinct(), 
                      by = "valve")

# Test dose effect on intercept and slope
intercept_model <- lm(intercept ~ dose, data = fit.dose)
slope_model <- lm(slope ~ dose, data = fit.dose)

cat("Effect of dose on intercept:\n")
print(summary(intercept_model))

cat("\nEffect of dose on slope:\n")
print(summary(slope_model))
```

```{r oav-results-table}
# Create summary table
kable(fit.dose, digits = 3, 
      caption = "Intercept and slope values by valve and dose")
```

## Temporal Visualization

```{r oav-temporal-plot, fig.cap="Total OAV over time by dose treatment"}
p5 <- ggplot(filtered_data_OAV, aes(x = elapsed_time, y = total_OAV, color = as.factor(dose))) +
  geom_point(size = 2, alpha = 0.7) +
  facet_wrap(~paste("Dose:", dose)) +
  labs(title = "Total OAV by Elapsed Time, Grouped by Dose",
       x = "Elapsed Time (minutes)",
       y = "Total OAV",
       color = "Dose") +
  theme_minimal() +
  theme(legend.position = "bottom")

print(p5)
```

## Machine vs Manual Application (OAV)

```{r oav-machine-manual}
# Filter for time = 0 comparison
filtered_data_4 <- OAV[OAV$group %in% c('Machine plot', 'No acid') & 
                         OAV$elapsed_time == 0, ]

if(nrow(filtered_data_4) > 0) {
  model_4 <- lm(total_OAV ~ group, data = filtered_data_4)
  
  # ANOVA
  anova_4 <- anova(model_4)
  print(anova_4)
  
  # Diagnostics
  shapiro_4 <- shapiro.test(model_4$residuals)
  filtered_data_4$res_abs <- abs(model_4$residuals)
  levene_4 <- anova(lm(res_abs ~ group, data = filtered_data_4))
  
  cat("Normality test p-value:", shapiro_4$p.value, "\n")
  cat("Homoscedasticity test p-value:", levene_4$`Pr(>F)`[1], "\n")
  
  # Means
  means_4 <- emmeans(model_4, ~ group)
  print(means_4)
} else {
  cat("No data available for time = 0 comparison\n")
}
```

```{r oav-application-plot, fig.cap="Initial OAV: Machine vs Manual application"}
if(exists("filtered_data_4") && nrow(filtered_data_4) > 0) {
  p6 <- ggplot(filtered_data_4, aes(x = group, y = total_OAV, color = group)) +
    geom_point(size = 3, alpha = 0.7) +
    stat_summary(fun = mean, geom = "crossbar", width = 0.5, color = "black") +
    labs(title = "Initial OAV: Application Method Comparison",
         x = "Application Method",
         y = "Total OAV",
         color = "Group") +
    theme_minimal() +
    theme(legend.position = "none")
  
  print(p6)
}
```

# Summary

## Key Findings

1. **VOC Emissions**: 
   - Dose effect: `r if(exists("model")) paste("p =", round(summary(model)$coefficients[2,4], 4)) else "Not calculated"`
   - Machine vs Manual: `r if(exists("anova_result_2")) paste("p =", round(anova_result_2$"Pr(>F)"[1], 4)) else "Not calculated"`

2. **NH₃ Emissions**:
   - Dose effect: `r if(exists("model_nh3")) paste("p =", round(summary(model_nh3)$coefficients[2,4], 4)) else "Not calculated"`
   - Machine vs Manual: `r if(exists("anova_3")) paste("p =", round(anova_3$"Pr(>F)"[1], 4)) else "Not calculated"`

3. **OAV Analysis**:
   - Dose effect on intercept: `r if(exists("intercept_model")) paste("p =", round(summary(intercept_model)$coefficients[2,4], 4)) else "Not calculated"`
   - Dose effect on slope: `r if(exists("slope_model")) paste("p =", round(summary(slope_model)$coefficients[2,4], 4)) else "Not calculated"`

## Conclusions

Based on the analysis, we can conclude that:

1. One incremental unit of acid significantly increased cumulative VOC emissions (after 119 h) by ~ 240 mg m^-2^
2. Machine application caused significantly higher cumulative VOC emissions (after 119 h) from untreated pig slurry (Machine plot) compared to manual application (0-DFC)
3. Acid did not have an effect on cumulative NH₃ emissions % TAN (after 119 h)
4. Machine application caused significantly higher cumulative NH₃ emissions (after 119 h) from untreated pig slurry (Machine plot) compared to manual application (0-DFC)
5. Acid did not have an effect on OAV, although I suspect it is because linear model may be not efficient in capturing the "trend". In fact, we have 0, 2.9, and 5.3 having the same OAV, and only 10.5 shows obvious higher OAV (at least at time 0)
6. No difference in OAV for Machine and Manual application of Raw pig slurry.

---

*Analysis completed on `r Sys.Date()`*
